Domestic cooking and likelihood

### 2022-07-01 inst/cook.Rmd
hyper3 analysis

\[
\frac{p_1+S}{p_1+S+p_2+p_3+p_4+p_5}\cdot
\frac{p_2  }{      p_2+p_3+p_4+p_5}\cdot
\frac{p_3  }{          p_3+p_4+p_5}\cdot
\frac{p_4  }{              p_4+p_5}\cdot
\frac{p_5  }{                  p_5}\cdot
\]

@@

>>>>

\[
\frac{p_1+S}{p_1+S+p_2+p_3+p_4+p_5}\cdot
\frac{p_2  }{      p_2+p_3+p_4+p_5}\cdot
\frac{p_3  }{          p_3+p_4+p_5}\cdot
\frac{p_4  }{              p_4+p_5}\cdot
\frac{p_5  }{                  p_5}
\]

@@

# Analysis with `hyper3` objects

We can take the "changes in cooking strength" augmented Plackett-Luce
likelihood functions above, and apply `hyper3` idiom to it.  Instead
of using a reified entity $S$ we have likelihood functions

\[
\frac{\lambda p_1}{\lambda p_1+p_2+p_3+p_4+p_5}\cdot
\frac{p_2  }{                  p_2+p_3+p_4+p_5}\cdot
\frac{p_3  }{                      p_3+p_4+p_5}\cdot
\frac{p_4  }{                          p_4+p_5}\cdot
\frac{p_5  }{                              p_5}
\]

\[
\frac{p_3  }{\lambda p_1+p_2+p_3+p_4+p_5}\cdot
\frac{\lambda p_1}{\lambda p_1+p_2+    p_4+p_5}\cdot
\frac{p_5  }{                  p_2+    p_4+p_5}\cdot
\frac{p_2  }{                  p_2+    p_4    }\cdot
\frac{p_4  }{                          p_4    }\cdot
\]

and then test $H_0\colon\lambda=1$.  The `hyper3` likelihood function
is mathematically the same as the reified `hyper2` BT likelihood
function provided $p_1+S=\lambda p_1$ (the likelihood functions differ
only if more than one player has increased strength); note that
$\lambda < 1$ is not possible with reified BT because it implies
negative $S$.  We will create a `hyper3` object for the post-training
observations:

```{r,hyper3post}
Hpost1 <- function(lambda){
   out <- hyper3()
   out[c(p1=lambda)] %<>% inc
   out[c(p1=lambda,p2=1,p3=1,p4=1,p5=1)] %<>% dec

   out[c(p2=1)] %<>% inc
   out[c(p2=1,p3=1,p4=1,p5=1)] %<>% dec

   out[c(p3=1)] %<>% inc
   out[c(p3=1,p4=1,p5=1)] %<>% dec

   out[c(p4=1)] %<>% inc
   out[c(p4=1,p5=1)] %<>% dec
   return(out)
}

Hpost2 <- function(lambda){
   out <- hyper3()
   out[c(p3=1)] %<>% inc
   out[c(p1=lambda,p2=1,p3=1,p4=1,p5=1)] %<>% dec

   out[c(p1=lambda)] %<>% inc
   out[c(p1=lambda,p2=1,p4=1,p5=1)] %<>% dec

   out[c(p5=1)] %<>% inc
   out[c(p2=1,p4=1,p5=1)] %<>% dec

   out[c(p2=1)] %<>% inc
   out[c(p2=1,p4=1)] %<>% dec
   return(out)
}
```

Now The likelihood functions for the "before"  observations simply add, so for example

```{r,showclever}
as.hyper3(Hbefore) + Hpost1(1.8888) + Hpost2(1.8888)
```

Then we can calculate a profile likelihood curve:

```{r,label=calculatesupportlambda,cache=TRUE}
H3 <- function(lambda){as.hyper3(Hbefore) + Hpost1(lambda) + Hpost2(lambda)}
supplam <- function(lambda){maxp(H3(lambda),n=1,give=TRUE)$likes}
lam <- exp(seq(from=log(0.8),to=log(100),len=50))
L <- sapply(lam,supplam)
```

```{r,label=plotsupp}
plot(lam,L-max(L),type="l",lwd=2)
abline(h=c(0,-2),col='gray')
abline(v=1,col='red')
text(0,-0.2,"lambda=1",col="red",pos=4)
```

We see that we can reject $H_0\colon\lambda=1$ but for this dataset
there does not appear to be a clear upper bound on $\lambda$.
